---
title: "11_lab"
author: "Aaron K"
date: "February 23, 2015"
output: html_document
---

```{r}
library(ggplot2)
library(dplyr)
require(tidyr)
require(broom)

df <-read.table("http://socserv.socsci.mcmaster.ca/jfox/Books/Applied-Regression-2E/datasets/UnitedNations.txt")
df$country <- rownames(df)
rownames(df) <- NULL
df <- tbl_df(df)
df
```

Plots
```{r}
ggplot(df, aes(lifeFemale, region)) +
  geom_point(colour = "red", position = "jitter") +
  geom_point(aes(lifeMale, region), colour = "blue", position = "jitter")
```

Class HW:
1. Describe the differences between Male and Female Life expectancy using stats.
2. Describe diff between M & F life Expect visually 
3. Describe diff between M & F life expect by region using stats
3b. what kind of variable is region?
4. Add region info to your plot.
5. challenge. redo the plot, but try adding the country names instead of points.

```{r}
# 1.
df %>% summarise(mean.life.female = mean(lifeFemale, na.rm=T), 
  mean.life.male = mean(lifeMale, na.rm=T))

# 2.
male.df <- df %>%
  filter(!is.na(lifeMale)) %>%
  select(-lifeFemale, life = lifeMale)
male.df$isMale <- 1
female.df <- df %>%
  filter(!is.na(lifeFemale)) %>%
  select(-lifeMale, life = lifeFemale)
female.df$isMale <- 0

df2 <- bind_rows(male.df, female.df)
ggplot(df2, aes(life)) +
  stat_bin()

```

Comparing Regions
1. create a dummy for european countries
2. create a dummy for non-euro countries
3. regress female life expect on the europe dummy. regress female life expectancy on the non-euro dummy. what do you notice?
4. regress female life expectancy on 'region'. interpret the results.
5. challenge. do euro and america have the same mean female life expect?

```{r}

df.2 <- df %>%
  # as.numeric forces it as 1 & 0 rather than T/F
  mutate(Europe=as.numeric(region=='Europe'),
         nonEurope=as.numeric(region!='Europe')) 

# 3
model1 <- lm(lifeFemale~Europe, df.2)
summary(model1)
tidy(model1)
model2 <- lm(lifeFemale~nonEurope, df.2)
summary(model2)
tidy(model2)
# 4 : it's the same thing if you focus on Europe
# dummy coding doesn't effect the slopes, random intercept (offsets)
model3 <- lm(lifeFemale~region, df.2)
tidy(model3)
# 5 : america's estimate is 3 years lower than europe (22vs19). 
# test if the difference is significant:
# it is not significantly different
levels(df$region)
df$region <- relevel(df$region, ref="Europe")
levels(df$region)
model4 <- lm(lifeFemale~region, df)
tidy(model4)
summary(model4)

```

If you put into R a variable with various factors (no dummy coding), then R will automatically create dummy codes and then use one of the factors as the baseline/reference. To check which one, look for which factor isn't present (use levels(df$region), to check all the factors of region). 

If you want to check if one level (factor) is signficantly different from another, change the reference factor : 
example : 
df$region  <- relevel(df$region, ref = "Europe")